Coursera Machine Learning Project
========================================================

Summary
-------
The goal of your project is to predict the manner how some exercises were conducted based on the data in the training dataset. My original plan was to test several methods available in the caret package on the training dataset, however I could conduct only the model fitting with *CART method* ("rpart"). Although I have cleaned the training dataset and removed some variables, the *random forest* method ("rf") was running too long, and the *gradient boosting* method ("gbm") generated an out of memory error.
With investigating the correlation between the variables I may get a  better and feasible model prediction.
With the rpart method I could achive a 54% accuracy which is fairly poor, I expect about 40% pass rate in the test dataset.

Background
----------
Using devices such as *Jawbone Up, Nike FuelBand*, and *Fitbit* it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify *how much* of a particular activity they do, but they rarely quantify *how well* they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har

The training data for this project are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har

Getting data
------------
The dataset is presented in comma separated format, it doesn't need any preformatting:
```{r}
library(caret)
training <- read.csv("pml-training.csv")
```
A quick review of the dataset's summary highlights that there is a lot of variable with missing or invalid data. The imputing method would be unreliable in case of a so high rate of missing data, so I simply discarded those variable and crated a dataset with the remaining variables.

Cleaning data
-------------
```{r}
variables <- c("user_name", "new_window", "num_window", "roll_belt", "pitch_belt",
               "yaw_belt", "total_accel_belt", "gyros_belt_x", "gyros_belt_y",
               "gyros_belt_z", "accel_belt_x", "accel_belt_y", "accel_belt_z",
               "magnet_belt_x", "magnet_belt_y", "magnet_belt_z", "roll_arm",
               "pitch_arm", "yaw_arm", "total_accel_arm", "gyros_arm_x",
               "gyros_arm_y", "gyros_arm_z", "accel_arm_x", "accel_arm_y",
               "accel_arm_z", "magnet_arm_x", "magnet_arm_y", "magnet_arm_z",
               "roll_dumbbell", "pitch_dumbbell", "yaw_dumbbell", "total_accel_dumbbell",
               "gyros_dumbbell_x", "gyros_dumbbell_y", "gyros_dumbbell_z",
               "accel_dumbbell_x", "accel_dumbbell_y", "accel_dumbbell_z", 
               "magnet_dumbbell_x", "magnet_dumbbell_y", "magnet_dumbbell_z",
               "roll_forearm", "pitch_forearm", "yaw_forearm", "total_accel_forearm",
               "gyros_forearm_x", "gyros_forearm_y", "gyros_forearm_z",
               "accel_forearm_x", "accel_forearm_y", "accel_forearm_z",
               "magnet_forearm_x",  "magnet_forearm_y", "magnet_forearm_z", "classe")
tr <- training[,variables]
```
Checking for near zero variance in variables show no possibility to remove variables:
```{r}
nsv <- nearZeroVar(tr, saveMetrics=TRUE)
nsv
```

The fitted model
----------------
The "glm" model is not possible, as the outcome variable has more that 2 levels. So my first choice was the rpart method, with the default cross validation method of the caret package:
```{r}
fit1 <- train(classe ~ ., data=tr, method="rpart")
```

Expected accuracy
-----------------

```{r}
fit1
```
Further investigation
---------------------
The analysis of the correlation between numeric variables (19 highly correlated variables!) shows a possibility for Principial Component Analysis, but unfortunately I had no time to do that.
```{r}
numVariables <- c("num_window", "roll_belt", "pitch_belt",
               "yaw_belt", "total_accel_belt", "gyros_belt_x", "gyros_belt_y",
               "gyros_belt_z", "accel_belt_x", "accel_belt_y", "accel_belt_z",
               "magnet_belt_x", "magnet_belt_y", "magnet_belt_z", "roll_arm",
               "pitch_arm", "yaw_arm", "total_accel_arm", "gyros_arm_x",
               "gyros_arm_y", "gyros_arm_z", "accel_arm_x", "accel_arm_y",
               "accel_arm_z", "magnet_arm_x", "magnet_arm_y", "magnet_arm_z",
               "roll_dumbbell", "pitch_dumbbell", "yaw_dumbbell", "total_accel_dumbbell",
               "gyros_dumbbell_x", "gyros_dumbbell_y", "gyros_dumbbell_z",
               "accel_dumbbell_x", "accel_dumbbell_y", "accel_dumbbell_z", 
               "magnet_dumbbell_x", "magnet_dumbbell_y", "magnet_dumbbell_z",
               "roll_forearm", "pitch_forearm", "yaw_forearm", "total_accel_forearm",
               "gyros_forearm_x", "gyros_forearm_y", "gyros_forearm_z",
               "accel_forearm_x", "accel_forearm_y", "accel_forearm_z",
               "magnet_forearm_x",  "magnet_forearm_y", "magnet_forearm_z")
M <- abs(cor(tr[,numVariables]))
diag(M) <- 0
View(M)
which(M>0.8, arr.ind=T)
```

Testing
-------